---
title: "Stock Analysis"
author: "Mac Faldet"
date: "February 4, 2019"
output: 
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Examples of Stock Analysis Practices

```{r libraries}
library(Quandl)
library(quantmod)
library(zoo)
```
quantmod provides a plethora of financial analysis tools helpful for importing and organizing stock ticker information.

The Op(), Hi(), Lo(), Cl(), Vo(), and Ad() functions can be used to extract the open, high, low, close, volume, and adjusted close column, respectively.

check out ?quantmod for more


```{r symbols}
#create an environment to keep our ticker information
stock_env <- new.env()

#identify the stocks for analysis
my_symbols <- c("AMZN", "BP", "C", "EEM", "IWM", "AAPL", "GOOG", "MSFT", "IBM", "GDP")
```
Stocks&ETFs| Amazon: AMZN, British Petroleum: BP, Citigroup: C, IShares MSCI Emerging MA: EEM, IShares Russell 2000 ETF: IWM
           | Apple: AAPL, Google: GOOG, Microsoft: MSFT, International Business Machines Corporation: IBM


```{r import}
#change the defaults for our getSymbols function to set the dates to pull
args("getSymbols.yahoo")
setDefaults(getSymbols.yahoo, from = "2014-01-01", to = "2019-01-01")

#view the defaults changed
getDefaults()
getDefaults(getSymbols.yahoo)

#load the ticker information into a specific environment
getSymbols(my_symbols, env = stock_env, auto.assign = TRUE)

#load US GDP information for comparison
#GDP <- Quandl("FRED/GDP", start_date="2014-01-01", end_date="2019-01-01", collapse = "weekly")
```
getSymbols gets ticker information from Yahoo Finance by default: getSymbols("AAA", src = "yahoo")
use the transform argument of Quandl to perform elementary calculations before collapse.

For example, getSymbols("CP", src = "yahoo") would load Canadian Pacific Railway data from the New York Stock Exchange. But getSymbols("CP", src = "FRED") would load Corporate Profits After Tax from the U.S. Bureau of Economic Analysis.

```{r stock Functions}
# View the high prices for all ticker series
head(Hi(stock_env$AMZN),20)

# Extract the Open price for AMZN
AMZN_open <- getPrice(stock_env$AMZN, prefer = "Open$")

# Look at January, 2016 using xts' ISO-8601 subsetting
AMZN_open["2016-01"]
```

For most intensive purposes, it is important to focus on the adjusted stock prices. Often Split and Dividend history affect how the overall open price is changed. Isolating the adjusted information will give us something to visualize more fairly.
```{r selected adjusted}
#extract volume column from each stock object and merge each list element into one object
adjusted <- do.call(merge, eapply(stock_env, Ad))

# Look at the first few lines of our adjusted prices
head(adjusted)

# Look at the column names of the stock object
colnames(adjusted)
```

lets generate some easy visualizations for our stock environment.
```{r portfolioPlots}
par(mfrow = c(3,1))
plot(adjusted$AMZN, main="Amazon")
plot(adjusted$BP, main="BP")
plot(adjusted$C, main="CitiGroup")

plot(adjusted$AAPL, main="Apple")
plot(adjusted$GOOG, main="Google")
plot(adjusted$MSFT, main="Microsoft")
```

now lets see a more aethetically pleasing visualizaiton given by ggplot2
```{r ggplot2}
library(ggplot2)


```


















## Case Study: Adjusting with only Close Prices

It is important to adjust our stock price visualizations for stock splits and dividends.

Stock splits can create large historical price changes even though they do not change the value of the company. So, you must adjust all pre-split prices in order to calculate historical returns correctly.

Similarly, you must adjust all pre-dividend prices. Dividends do reduce the company's value by the amount of the dividend payment, but the investor's return isn't affected because they receive the offsetting dividend payment.

```{r adjustPortfolio}
adjusted_env <- do.call(merge, eapply(stock_env, adjustOHLC))

head(adjusted_env)
```

adjustOHLC() only works for OHLC data. It will not work if you only have close prices, and it does not return any of the split or dividend data it uses.

You need the dates and values for each split and dividend to adjust a non-OHLC price series, or if you simply want to analyze the raw split and dividend data.

You can download the split and dividend data from Yahoo Finance using the quantmod functions getSplits() and getDividends(), respectively. The historical dividend data from Yahoo Finance is adjusted for splits. If you want to download unadjusted dividend data, you need to set split.adjust = FALSE in your call to getDividends().

```{r AMZNonlyCloseAdjustments}
# Download AAPL split data
splits <- getSplits(AMZN)

# Download AAPL dividend data
dividends <- getDividends(AMZN)

# Look at the first few rows of dividends
head(dividends)

# Calculate split and dividend adjustment ratios
ratios <- adjRatios(splits = splits, dividends = raw_dividends, close = Cl(AMZN))

# Calculate adjusted close for AAPL
aapl_adjusted <- Cl(AMZN) * ratios[, "Split"] * ratios[, "Div"]

# Look at first few rows of Yahoo adjusted close
head(Ad(AMZN))

# Look at first few rows of aapl_adjusted
head(aapl_adjusted)
```



## Case Study: Files

Here's an example of creating, uploading, and downloading files containing ticker symbols
```{r upload}
# Save symbol lookup table
saveSymbolLookup(file = "AMZN.csv")

# Set default source for CP to "yahoo"
setSymbolLookup(AMZN = "yahoo")

# Verify the default source is "yahoo"
getSymbolLookup("AMZN")

# Load symbol lookup table
loadSymbolLookup(file = "AMZN.csv")
```

```{r download}
# Load AMZN.csv
getSymbols("AMZN", src = "csv")

# Look at AMZN structure
str(AMZN)
```
or
```{r zooDownload}
# Import AMZN.csv using read.zoo
amzn_zoo <- read.zoo("AMZN.csv", sep = ",", header = TRUE)

# Convert to xts
amzn_xts <- as.xts(amzn_zoo)

# Look at the first few rows of amzn_xts
head(amzn_xts)
```
this download using read.zoo(), provided by the zoo package, wraps the read.table() function with the as.xts() funtion.



## Case Study: Checking for Weirdness

We will use open stock prices from AMZN to evaluate different methods of filling NA's
```{r}
# fill NA using last observation carried forward
locf <- na.locf(Op(AMZN))

# fill NA using linear interpolation
approx <- na.approx(Op(AMZN))

# fill NA using spline interpolation
spline <- na.spline(Op(AMZN))

# merge into one object
na_filled <- merge(locf, approx, spline)

# plot combined object
plot(na_filled, col = c("black", "red", "green"))
```



## Case Study: Time Stamps

Sometimes two series have the same periodicy, but use different conventions to represent a timestamp. For example, monthly series may be timestamped with the first or last date of the month. The different timestamp convention can cause many NA when series are merged. The yearmon class from the zoo package helps solve this problem.

Lets aggregate the FRED daily Fed Funds rate (DFF) to a monthly periodicy and merge it with the FRED monthly Fed Funds rate (FEDFUNDS).The DFF aggregate will be timestamped with the last row of the month, while FEDFUNDS is timestamped with the first day of the month.

```{r timeStamp merge}
# Download FEDFUNDS and DFF data
getSymbols(c("FEDFUNDS", "DFF"), src = "FRED")

# Aggregate DFF to monthly
monthly_fedfunds <- apply.monthly(DFF, mean)

# Convert index to yearmon
index(monthly_fedfunds) <- as.yearmon(index(monthly_fedfunds))

# Merge FEDFUNDS with the monthly aggregate
merged_fedfunds <- merge(FEDFUNDS, monthly_fedfunds)

# Look at the first few rows of the merged object
head(merged_fedfunds)
```

A more manual method of merging daily timestamped data with monthly is as follows:

First you merge the lower-frequency data with the aggregate data, then use na.locf() to fill the NA forward (or backward, using fromLast = TRUE). Then you can subset the result using the index of the object with the representation you prefer.

Reminder: our workspace contains FEDFUNDS, monthly_fedfunds (the result of apply.monthly(DFF, mean)), and merged_fedfunds (the result of merge(FEDFUNDS, monthly_fedfunds) where the monthly_fedfunds index is a Date). Note the NA values in monthly_fedfunds.

```{r manual timestampMerge}
# Fill NA forward
merged_fedfunds_locf <- na.locf(merged_fedfunds)

# Extract index values containing last day of month
aligned_last_day <- merged_fedfunds_locf[index(monthly_fedfunds)]

# Fill NA backward
merged_fedfunds_locb <- na.locf(merged_fedfunds, fromLast = TRUE)

# Extract index values containing first day of month
aligned_first_day <- merged_fedfunds_locb[index(FEDFUNDS)]
```

A general aggregation technique to aggregate daily data to weekly, but with weeks ending on Wednesdays. This is often done in stock market research to avoid intra-week seasonality.

You can supply your own end points to period.apply() (versus using endpoints()). Recall endpoints() returns locations of the last observation in each period specified by the on argument. The first and last elements of the result are always zero and the total number of observations, respectively. The end points you pass to period.apply() must follow this convention.

This exercise will use the daily Fed Funds data (DFF) from prior exercises.

```{r aggregateWeekly}
# Extract index weekdays
index_weekdays <- .indexwday(DFF)

# Find locations of Wednesdays
wednesdays <- which(index_weekdays == 3)

# Create custom end points
end_points <- c(0, wednesdays, nrow(DFF))

# Calculate weekly mean using custom end points
weekly_mean <- period.apply(DFF, end_points, mean)
```

Sometimes financial data is even intraday, and contains seconds. Here is how to regulate intraday timestamps

Note: merging with object with merge() will result in a timeseries that takes the same timezone as the first object.

```{r intraday}
datetimes <- seq(from = as.POSIXct("2016-01-16 08:00"),
                 to = as.POSIXct("2016-01-17 18:00"),
                 by = "1 min")

regular_xts <- xts(dc_trades, order.by = datetimes)
merged_xts <- merge(dc_trades, regular_xts)

head(merged_xts)

# All observations should be NA
all(is.na(merged_xts["2016-01-16 19:00/2016-01-17 07:00"]))

# xts time-of-day subsetting
merged_trade_day <- merged_xts["T08:00/T18:00"]

# Now there are no observations
nrow(merged_trade_day["2016-91-16 19:00/2016-01-17 07:00"])

# Fill missing values by trading times

# split() data into list of non-overlapping chunks
trade_day_list <- split(merged_trade_day, "days")

#lapply() a function to each chunk (list element)
filled_trade_day_list <- lapply(trade_day_list, na.locf)

# Combine list of chunks using do.call() and rbind()
filled_trade_day <- do.call(rbind, filled_trade_day_list)
```
or
```{r intraday2}
# Create a regular date-time sequence
regular_index <- seq(as.POSIXct("2010-01-04 09:00"), as.POSIXct("2010-01-08 16:00"), by = "30 min")

# Create a zero-width xts object
regular_xts <- xts(x = NULL, order.by = regular_index)

# Merge irregular_xts and regular_xts, filling NA with their previous value
merged_xts <- merge(irregular_xts, regular_xts, fill = na.locf)

# Subset to trading day (9AM - 4PM)
trade_day <- merged_xts["T09:00/T16:00"]
```

The previous exercise carried the last observation of the prior day forward into the first observation of the following day. This exercise will show you how to fill missing values by trading day, without using the prior day's final value.

You will use the same split-lapply-rbind paradigm from the Introduction to xts and zoo course with DataCamp. For reference, the pattern is below.

x_split <- split(x, f = "months")
x_list <- lapply(x_split, cummax)
x_list_rbind <- do.call(rbind, x_list)

Recall that the do.call(rbind, ...) syntax allows you to pass a list of objects to rbind() instead of having to type all their names.

Your workspace has a trade_day object that contains the regular series from the previous exercise, but without any NA filled in.

```{r intraday2 cont}
# Split trade_day into days
daily_list <- split(trade_day , f = "days")

# Use lapply to call na.locf for each day in daily_list
daily_filled <- lapply(daily_list, FUN = na.locf)

# Use do.call to rbind the results
filled_by_trade_day <- do.call(rbind, daily_filled)
```

Intraday data can be huge, with hundreds of thousands of observations per day, millions per month, and hundreds of millions per year. These data sets often need to be aggregated before you can work with them.

You learned how to aggregate daily data in the Introduction to xts and zoo course. This exercise will use to.period() to aggregate intraday data to an OHLC series. You often need to specify both period and k arguments to aggregate intraday data.

The intraday_xts object contains one trading day of random data.

```{r aggregate intraday}
# Convert raw prices to 5-second prices
xts_5sec <- to.period(intraday_xts, period = "seconds", k = 5)

# Convert raw prices to 10-minute prices
xts_10min <- to.period(intraday_xts, period = "minutes", k = 10)

# Convert raw prices to 1-hour prices
xts_1hour <- to.period(intraday_xts, period = "hours", k = 1)
```




##Other Helpful Examples

The getPrice() function in the quantmod package can extract any column by name by using the prefer argument. It can also extract columns for a specific instrument by using the symbol argument, which is useful when an object contains several instruments with the same price type.
```{r usingQuandl}
# Download CME data for CL and BZ as an xts object
qtr_oil_price <- Quandl(code = c("CME/CLH2016", "CME/BZH2016"), type = "xts", collapse = "quarterly")

# Look at the column names of the oil_data object
colnames(qtr_oil_price)

# View the high prices for both series
Hi(qtr_oil_price)

# Extract the Open price for CLH2016
cl_open <- getPrice(qtr_oil_price, symbol = "CLH2016", prefer = "Open$")

# Look at January, 2016 using xts' ISO-8601 subsetting
cl_open["2016-01"]

# Download quarterly CL and BZ returns
qtr_return <- Quandl(quandl_codes, type = "xts", collapse = "quarterly", transform = "rdiff")

# View the settle price returns for both series
getPrice(qtr_return, prefer = "Settle")
```


When tracking financial metrics, there are often times you are working with an irregular time series. To work around missing dates for holidays or other reasons, use the code:
```{r regulating time series}
# Extract the start date of the series
end_date <- start(irregular_xts)

# Extract the end date of the series
end_date <- end(irregular_xts)

# Create a regular date sequence
regular_index <- seq(from = start_date,
                     to = end_date,
                     by = "day")

# Create a zero-width xts object
regular_xts <- xts(irregular_xts, order.by = regular_index)

# Merge irregular_xts and regular_xts
merged_xts <- merge(regular_xts, irregular_xts)

# Look at the first few rows of merged_xts
head(merged_xts)

# Use the fill argument to fill NA with their previous value
merged_filled_xts <- merge(regular_xts, irregular_xts, fill = na.locf)

# Look at the first few rows of merged_filled_xts
head(merged_filled_xts)
```


Here's an example of using 'eapply' to perform a task on a list of data, such as our stock environment.
```{r eapply ex}
# Call head on each object in data_env using eapply and merge all the list elements into one xts object
data_merged <- do.call(merge, eapply(stock_env, head))

# Ensure the columns are ordered:open, high, low, close
data_ohlc <- OHLC(data_merged) 
```

Here's an example of using Ticker symbols with syntactly incorrect names
```{r nonSyntact exOne}
# Load BRK-A data
getSymbols("BRK-A")

# Use backticks and head() to look at the loaded data
head(`BRK-A`)

# Use get() to assign the BRK-A data to an object named BRK.A
BRK.A <- get("BRK-A")
```
or
```{r nonSyntact exTwo}
# Create BRK.A object
BRK.A <- getSymbols("BRK-A", auto.assign = FALSE)

# Create col_names object with the column names of BRK.A
col_names <- colnames(BRK.A)

# Set BRK.A column names to syntactically valid names
colnames(BRK.A) <- make.names(col_names)
```
or
```{r nonSyntact exThree}
# Set name for BRK-A to BRK.A
setSymbolLookup(BRK.A = list(name = "BRK-A"))

# Set name for T (AT&T) to ATT
setSymbolLookup(ATT = list(name = "T"))

# Load BRK.A and ATT data
getSymbols(c("BRK.A", "ATT"))
```

